---
title: "Excercise 2"
output: pdf_document
author: "Ming Martin Liu"
date: "1/13/2021"
---
# Excercise 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, include = FALSE}
#library
library(expss)
library(pracma)
library(knitr)
library(MASS)
library(pander)
library(ggplot2)
```

```{r, echo=FALSE}
#params
N = rep(0,10)
D = rep(0,10)
V = rep(0,10)
dt = c(1:10)
t = seq(0, 9000, by=10)
Q = rep(0,901)
c = 750/pi
col = seq(1, length(N), by = 1)
g = 32.2 #ft/s^2

```

```{r, echo=FALSE}
#function
f <- function(t){
  c*(1-cos((pi*t)/4500))
}

Q = f(t)

A = trapz(t,Q)

B = trapzfun(f, 0, 9000, maxit = 50, tol = 10)

plot(t,Q,type='l', 
     col='blue',
     main = 'Graph of flowrate versus Time', 
     xlab = 'Time (s)', 
     ylab = expression('Q ('*ft^3*'/s)'),
     xaxp = c(0,9000,9),
     yaxp = c(0,500,5)
)
```


\begin{center}
Figure 1: Graph of Flow vs Time.
\end{center}
```{r, echo=FALSE}
S_total = rep(0,10)
for(i in 1:10){
  t1 = seq(0,9000, by = dt[i])
  Q = f(t1)
  N[i] = floor(9000/(dt[i]))
  V[i] = trapz(t1,Q)
  S = rep(0,length(t1))
  for (j in 1:length(t1)){
    S[j+1] = S[j] + (Q[j] + Q[j+1])*0.5*dt[i]
    S_total[i] = S[length(t1)]
  }
}

df = data.frame(col,N,V)

```
## Checking for accuracy

This process is conducted by checking that the volume of Storage in the reservoir remains constant with a varying number of iteratious

\begin{center}
Table 1 shows the Volume of Flow, Q with respect to the number of iterations using the Trapezoid formula.
\end{center}
```{r comment='', echo=FALSE, results='asis'}
 knitr::kable(
  df,
  col.names = c(' ', 'Number of Iterations', 'Volume ($ft^3$)'),
  align = "lcc"
)
```

As shown in the table, the difference between a timestep of 1 and a timestep of 10 is negligible, hence a timestep of 10 is used as it reduces runtime while retaining accuracy.

# Excercise 2
```{r}
#exercise 2
hmax = 6 #ft
R = 2 #ft
CD = 0.65
Area = 2148590/6 #assume this constant
#df1 = data.frame(col,N,S_total)
```

## Case 1: Reservoir filling

### Checking for accuracy using Number of Iterations


In order to find Storage at some time t, we use data from the previous exercise, namely the flowrate, Q. In addition in order to find an appropriate timestep, we use the trapezoidal method with decreasing timesteps until we observe negligible difference.

This is done by observing maximum height and storage with respect to the Number of Iterations, N.


```{r}
# time to fill is fixed at 9000s
# Reference to code from ex.1
S_max = 0
H_max = 0

for(i in 1:10){
  t1 = seq(0,9000, by = dt[i])
  Q = f(t1)
  N[i] = floor(9000/(dt[i]) + 1)
  S_fill = rep(0,length(t1))
  H_fill = rep(0,length(t1))
  for (j in 1:length(t1)){
    S_fill[j+1] = S_fill[j] + (Q[j] + Q[j+1])*0.5*dt[i]
  }
  S_max[i] = S_fill[length(t1)]
  H_max[i] = S_max[i]/Area
}

df2 = data.frame(col,N,S_max,H_max)
```

\begin{center}
Table 2 shows the Max Height and Storage with respect to the Number of Iterations
\end{center}
```{r comment='', echo=FALSE, results='asis'}
 knitr::kable(
  df2,
  col.names = c(' ', 'Number of Iterations', 'Max Storage ($ft^3$)','Max Height (ft)'),
  align = "lcc"
)
```
### Plotting Graphs for Filling Reservoir

As can be seen in Table 2, despite the lower Number of Iterations, the Max Storage and Height remain constant. Hence 900 iterations, which is equivalent to a timestep of 10 seconds, is acceptable.

```{r}
S_fill = 0 # Creating empty vectors
H_fill = 0
for(i in 1:900){
  S_fill[i+1] = S_fill[i] + (Q[i] + Q[i+1])*0.5*10
} 

H_fill = S_fill/360000
S_thousands = S_fill/1000
```

Next we plot Storage vs time and height vs time


```{r}
plot(t,S_thousands,type='l', 
     col='blue',
     main = 'Graph of Storage versus Time', 
     xlab = 'Time (s)', 
     ylab = expression('S ' * x10^3*'('*ft^3*')'),
     xaxp = c(0,9000,9),
)
```

\begin{center}
Figure 2: Graph of Filling Reservoir, Storage vs Time.
\end{center}

```{r}
plot(t,H_fill,type='l', 
     col='blue',
     main = 'Graph of Height versus Time', 
     xlab = 'Time (s)', 
     ylab = expression('Height (ft)'),     
     xaxp = c(0,9000,9),
)
```

\begin{center}
Figure 3: Graph of Filling Reservoir, Height vs Time.
\end{center}


## Case 2: Reservoir emptying

Our design choice is for outlfow to occur through a pipe at base of reservoir, with the entry to pipe being the orifice.

```{r}
#Init variables

d = 2 #orifice diameter
a = pi
const = CD*a*((2*g)^0.5)
Q_empty = 0
t2empty = 0
```

### Checking for accuracy using time steps

In this code chunk we are trying to identify which time step is the most efficient, simply put we want something that is small enough to be accurate but large enough to reduce runtime. Ideally, we will achieve similar results as before, where time step will be negligible. Meaning that the time to empty is similar across all time steps.


```{r}
for (i in 1:10){
  S_empty = 2148590
  H_empty = 6
  n=1
  while(S_empty[n]>0){
    S_empty[n+1] = S_empty[n] - const*dt[i]*(H_empty[n])^0.5
    H_empty[n+1] = S_empty[n+1]/360000
    n=n+1
  }
  S_empty=head(1:length(S_empty), -1)
  t2empty[i]=dt[i]*length(S_empty)
}
df3 = data.frame(col,dt,t2empty)


```

```{r comment='', echo=FALSE, results='asis'}
 knitr::kable(
  df3,
  col.names = c(' ', 'Time Step (s)', 'Time To Empty(s)'),
  align = "lcc"
)
```
 
As can be seen in Table 3, the time to empty does vary based on the timestep used. This is most likely because as the timestep gets larger, a larger part of the Storage is being cut off due to the timestep. Hence for this section a time step of 1 second will be used to achieve the most accurate results.

### Checking for consistent Flow

Next, with a timestep of 1, the area under the flowrate versus time graph is calculated to see if the total storage still remains as 2148590.
```{r}

## test run with timestep of 1 second
S_empty = 2148590
H_empty = 6
t_empty = 0
Q_empty = 0
n=1
while(S_empty[n]>0){
    S_empty[n+1] = S_empty[n] - const*10*(H_empty[n])^0.5
    H_empty[n+1] = S_empty[n+1]/360000
    t_empty[n+1] = t_empty[n] + 1
    n=n+1
}
Q_empty = 0.65*a*(2*g*H_empty)^0.5
sum=0
limit = length(S_empty)-2
for (j in 1:limit){
  sum = sum + (Q_empty[j]+Q_empty[j+1])*10*0.5
}


#Converting H and T for better axis

S_empty = S_empty/10^4
t_empty = t_empty/10^3
```

Based on the calculations, the area under graph is 2148570, which is approximately off by 0.00093%. This difference can be taken to be negligible compared to the original.

### Plotting Graphs for Emptying Reservoir

Using a time step of 1 second, the graphs of Flowrate vs time, Storage vs time and Height vs time are plotted below

```{r}
plot(t_empty,Q_empty,type='l', 
     col='blue',
     main = 'Graph of Outflow versus Time', 
     xlab = expression('Time ' * x10^3*'(s)'), 
     ylab = expression('Q '*'('*ft^3*'/s)'),
)
```


\begin{center}
Figure 4: Graph of Emptying Reservoir, Flowrate vs Time.
\end{center}


```{r}

plot(t_empty,S_empty,type='l', 
     col='blue',
     main = 'Graph of Storage emptying versus Time', 
     xlab = expression('Time ' * x10^3*'(s)'), 
     ylab = expression('S ' * x10^4*'('*ft^3*')'),

)
```

\begin{center}
Figure 5: Graph of Emptying Reservoir, Storage vs Time.
\end{center}

```{r}
t_try=seq(0, 9000, by=1)
plot(t_empty,H_empty,type='l', 
     col='blue',
     main = 'Graph of Height of reservoir level versus Time', 
     xlab = expression('Time ' * x10^3*'(s)'), 
     ylab = expression('H (ft)'),

)
```


\begin{center}
Figure 6: Graph of Emptying Reservoir, Height vs Time.
\end{center}

